---
title: "Minifish markdown"
author: "Orlando Bruno Giorgetti"
date: "2/6/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(reshape2)
library(ggplot2)
library(Biostrings)
library(dplyr)
library(igraph)
pc.chart = function(pc.list,i,j,k=0)sapply(pc.list[ sapply(pc.list,length)>=(i+k)],function(x)sapply(pc.list[ sapply(pc.list,length)>=(i+k)],function(y)length(intersect(x[sample((1+k):(i+k),j)],y[sample((1+k):(i+k),j)]))))
pc.chart.full = function(pc.list,j,k=0)sapply(pc.list[ sapply(pc.list,length)>=(k+j)],function(x)sapply(pc.list[ sapply(pc.list,length)>=(j)],function(y)length(intersect(x[sample((1+k):length(x),j)],y[sample((1+k):length(y),j)]))))
matrix.builder <- function(distancia){
    matriz = as.matrix(distancia)
    matriz[upper.tri(matriz)] <- NA
    diag(matriz) = NA
    return(matriz)  
}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```


```{r Minifish statistics, echo = FALSE, message=FALSE, results='hide',warning=FALSE,fig.keep='all'}
load(file = "MF.list")
MF.x.list = MF.list
MF.fish.names = c("Fish 2","Fish 3","Fish 4","Fish 5")
MF.gene.names = c("TCRa","TCRb","TCRg","TCRd","IgH")
names(MF.x.list) = MF.gene.names
for (i in 1:length(MF.x.list))names(MF.x.list[[i]]) = MF.fish.names
MF.Ja.alleles.names = c(1,6,10,12,21,30,31,34,39,42,43,45,47,48,50,51,55,60)
MF.x.df = lapply(MF.x.list,function(x)do.call(rbind,x))
MF.x.df = lapply(MF.x.df,function(x)cbind(x,fish=sub("\\..*","",rownames(x))))
```

# Minifish clones

Number of clones

```{r Clones,echo = FALSE}
MF.df.clones = data.frame(sapply(MF.x.list,function(x)sapply(x,nrow)),row.names = MF.fish.names)
colnames(MF.df.clones) = MF.gene.names
kable(MF.df.clones)
MF.df.clones.melt = melt(MF.df.clones,measure.vars = colnames(MF.df.clones))
MF.df.clones.melt$fish = c(MF.fish.names)
ggplot(data = MF.df.clones.melt[1:16,]) + geom_bar( mapping = aes (x = fish,y =value,fill = variable), width = 0.5,stat = "identity") + ylab ("clone count" ) + ggtitle ("clone count")
ggplot(data = MF.df.clones.melt[17:20,]) + geom_bar( mapping = aes (x = fish,y =value,fill = variable), width = 0.5,stat = "identity") + ylab ("clone count" ) + ggtitle ("clone count")

```


# Minifish cDNA molecules per fish

Number of UMIs matched to a receptor sequence (not unique)

```{r Molecules,echo = FALSE}
MF.df.molecules = data.frame(sapply(MF.x.list,function(x)sapply(x,function(y)sum(y$Umi.count))),row.names = MF.fish.names)

colnames(MF.df.molecules) = MF.gene.names
kable(MF.df.molecules)
MF.df.molecules.melt = melt(MF.df.molecules,measure.vars = colnames(MF.df.molecules))
MF.df.molecules.melt$fish = c(MF.fish.names)
ggplot(data = MF.df.molecules.melt[1:16,]) + geom_bar( mapping = aes (x = fish,y =value,fill = variable), width = 0.5,stat = "identity") + ylab ("UMI/cDNA molecules count" ) + ggtitle ("TCR UMI count")
ggplot(data = MF.df.molecules.melt[17:20,]) + geom_bar( mapping = aes (x = fish,y =value,fill = variable), width = 0.5,stat = "identity")+ ylab ("UMI/cDNA molecules count" ) + ggtitle ("IgH UMI count")

```


# Structure of the TCR alpha beta repertoire



```{r Structure, echo = FALSE}
ggplot(MF.x.list$TCRa$`Fish 5`) + geom_point(mapping = aes(y=Umi.proportion, x = 1:length(Umi.proportion),color = factor(ceiling(cumsum(MF.x.list$TCRa$`Fish 5`$Umi.proportion)/0.2)))) + labs (color = "Quintile") + xlab("Sequence rank") + ylab("Umi count proportion") + ggtitle("TCR alpha rank-frequency plot")
```






```{r Structure beta, echo = FALSE}

ggplot(MF.x.list$TCRb$`Fish 5`) + geom_point(mapping = aes(y=Umi.proportion, x = 1:length(Umi.proportion),color = factor(ceiling(cumsum(MF.x.list$TCRb$`Fish 5`$Umi.proportion)/0.2)))) + labs (color = "Quintile") + xlab("Sequence rank") + ylab("Umi count proportion") + ggtitle("TCR beta rank-frequency plot")
```
Approximately half of the Molecules(Cells) in TCR alpha are distributed among the top 250 clones, 
while the remaining 3500 molecules are distributed across 2500 clones

Approximately half of the Molecules(Cells) in TCR beta are distributed among the top 500 clones, 
while the remaining 3500 molecules are distributed across 2100 clones

# Structure of the TCR gamma repertoire


Approximately half of the Molecules(Cells) in TCR beta are distributed among the top 500 clones, 
while the remaining 3500 molecules are distributed across 2100 clones


```{r Structure gamma, echo = FALSE}

ggplot(MF.x.list$TCRg$`Fish 5`) + geom_point(mapping = aes(y=Umi.proportion, x = 1:length(Umi.proportion),color = factor(ceiling(cumsum(MF.x.list$TCRg$`Fish 5`$Umi.proportion)/0.2)))) + labs (color = "Quintile") + xlab("Sequence rank") + ylab("Umi count proportion") + ggtitle("TCR gamma rank-frequency plot")
```

```{r Structure delta, echo = FALSE}

ggplot(MF.x.list$TCRd$`Fish 5`) + geom_point(mapping = aes(y=Umi.proportion, x = 1:length(Umi.proportion),color = factor(ceiling(cumsum(MF.x.list$TCRd$`Fish 5`$Umi.proportion)/0.2)))) + labs (color = "Quintile") + xlab("Sequence rank") + ylab("Umi count proportion") + ggtitle("TCR delta rank-frequency plot")
```

```{r Structure IgH, echo = FALSE}

ggplot(MF.x.list$IgH$`Fish 5`) + geom_point(mapping = aes(y=Umi.proportion, x = 1:length(Umi.proportion),color = factor(ceiling(cumsum(MF.x.list$IgH$`Fish 5`$Umi.proportion)/0.2)))) + labs (color = "Quintile") + xlab("Sequence rank") + ylab("Umi count proportion") + ggtitle("IgH rank-frequency plot")
```

# Singletons
How many of the cDNA molecules are singletons (only one cDNA found for that receptor)

```{r Singletons, echo = FALSE}
MF.df.singletons = data.frame(sapply(MF.x.list,function(x)sapply(x,function(y)mean(y$Umi.count==1))),row.names = MF.fish.names)
colnames(MF.df.singletons) = MF.gene.names
kable(round(MF.df.singletons,3))
```

# Publicity


```{r Publicity table, echo = FALSE,fig.width= 12,fig.height=10}

  
pub.df = data.frame(publicity = 1:4,sapply(1:5,function(i)table(do.call(rbind,MF.x.list[[i]])$pub)/1:4))
colnames(pub.df)[2:6] = MF.gene.names
kable(pub.df)
```

```{r Publicity, echo = FALSE,fig.width= 12,fig.height=10}
{
a=prop.table(table(MF.x.df$TCRa[!duplicated(MF.x.df$TCRa$pc.nt) & MF.x.df$TCRa$fish != "Fish 1",]$pub))
b=prop.table(table(MF.x.df$TCRb[!duplicated(MF.x.df$TCRb$pc.nt) & MF.x.df$TCRb$fish != "Fish 1",]$pub))

pub_ab=data.frame(proportion = c(a,b),publicity = rep(1:4,2), gene = rep(c("TCRa","TCRb"),each =4))

ggplot(data = pub_ab) + 
  geom_point( mapping = aes (x = publicity,y =proportion,shape = gene),size = 5) +
  geom_line( mapping = aes (x = publicity,y =proportion,color = gene)) +
   theme(text = element_text(size=24) )

}
```

```{r Publicity 2, echo = FALSE,fig.width= 12,fig.height=10}

ggplot(data = pub_ab,mapping =  aes (x = log(publicity),y =log(proportion),shape = gene,color = gene)) + 
  geom_point(color = "black",size = 7) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_classic() +
   theme(text = element_text(size=10),
       axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
       axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
       )



g=prop.table(table(MF.x.df$TCRg[!duplicated(MF.x.df$TCRg$pc.nt) & MF.x.df$TCRg$fish != "Fish 1",]$pub))
d=prop.table(table(MF.x.df$TCRd[!duplicated(MF.x.df$TCRd$pc.nt) & MF.x.df$TCRd$fish != "Fish 1",]$pub))

pub_gd=data.frame(proportion = c(g,d),publicity = rep(1:4,2), gene = rep(c("TCRg","TCRd"),each =4))

pub_gd$gene = factor(pub_gd$gene, levels = c("TCRg","TCRd"))
```

```{r Publicity 3, echo = FALSE,fig.width= 12,fig.height=10}

ggplot(data = pub_gd) + 
  geom_point( mapping = aes (x = publicity,y =proportion,shape = gene),size = 5) +
  geom_line( mapping = aes (x = publicity,y =proportion,color = gene)) +
  scale_shape_manual(values=c(16,17)) + 
  scale_color_manual(values=c( gg_color_hue(4)[2],gg_color_hue(4)[4])) +
  theme(text = element_text(size=24) )
```

```{r Publicity 4, echo = FALSE,fig.width= 12,fig.height=10}

ggplot(data = pub_gd,mapping =  aes (x = log(publicity),y =log(proportion),shape = gene,color = gene)) + 
  geom_point(color = "black",size = 7) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_shape_manual(values=c(16,17)) + 
  scale_color_manual(values=c( gg_color_hue(4)[2],gg_color_hue(4)[4])) +
   theme_classic() +
 theme(text = element_text(size=24),
       axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
       axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
       )


#ggplot(data = MF.x.df$TCRb) + geom_bar( mapping = aes (x = pub),position = "dodge") + facet_wrap( ~fish)  + labs(title="TCR beta publicity")
#kable(table(MF.x.df$TCRa$pub))
```

```{r Publicity 5, echo = FALSE,fig.width= 12,fig.height=10}
{
g=make_full_graph(4, directed = FALSE, loops = FALSE)
V(g)$label = paste0("Fish_",2:5)
sh=as.numeric(matrix.builder(pc.chart(lapply(MF.x.list$TCRa,function(x)x$pc.nt),500,500)))
sh = sh[!is.na(sh)]
sh2=as.numeric(matrix.builder(pc.chart(lapply(MF.x.list$TCRb,function(x)x$pc.nt),500,500)))
sh2 = sh2[!is.na(sh2)]
#E(g)$width = sh/10

E(g)$width = sh/6
E(g)$color =gg_color_hue(2)[1]
g3 = add_edges(g,c(1,2, 1,3, 1,4, 2,3, 2,4, 3,4))
E(g3)$width[7:12] = sh2/6
E(g3)$color[7:12] = gg_color_hue(2)[2]
E(g3)$curved = rep(c(-0.15,0.15),each=6)
lkk = layout.kamada.kawai(g3)
plot(g3,vertex.frame.color = "grey", vertex.color  = "grey",vertex.size = 20, vertex.label.color = "black",layout = lkk)
}
```

```{r Publicity 6, echo = FALSE,fig.width= 12,fig.height=10}


pub.mean = data.frame(gene = c("TCRa","TCRb"),mean = c(mean(sh,na.rm = TRUE),mean(sh2,na.rm = TRUE)))

sh.names = c("2_3","2_4","2_5","3_4","3_5","4_5")
sh.a.df = data.frame(sh,sh.names)
sh.b.df = data.frame(sh2,sh.names)

sh.table.1= cbind(sh.a.df, x = 1)
sh.table.2= cbind(sh.b.df, x = 2)
colnames(sh.table.2)[1] = "sh"
sh.table = rbind(sh.table.1,sh.table.2)
```

```{r Publicity 7, echo = FALSE,fig.width= 10,fig.height=12}

ggplot(data = pub.mean) + geom_bar(mapping = aes (x = gene , y=mean, fill = gene) , width = 0.3,stat = "identity") + 
  ggtitle ("Mean pairwise overlap top 500 clones") + 
  geom_point(data = sh.a.df,mapping = aes (x = 1, y = sh)) +
  geom_line (data = sh.table,mapping = aes(x = x, y = sh,group = sh.names)) +
  geom_text(data = sh.a.df,mapping = aes (x = rep(0.8,6), y = sh+c(-0,+0,rep(0,4)), label = sh.names,fontface = "italic"),size = 8) +
  geom_point(data = sh.b.df,mapping = aes (x = 2, y = sh2)) +
  geom_text(data = sh.b.df,mapping = aes (x = c(1.8,2.2,2.2,2.2,2.2,2.2), y = sh2+c(-0,+0,-1,rep(0,3)), label = sh.names,fontface = "italic"),size = 8) +
  theme(text = element_text(size=20),
        axis.text=element_text(size=20)) 

```

```{r Publicity 8, echo = FALSE,fig.width= 10,fig.height=10}

{
g=make_full_graph(4, directed = FALSE, loops = FALSE)
V(g)$label = paste0("Fish_",2:5)
sh=as.numeric(matrix.builder(pc.chart(lapply(MF.x.list$TCRg,function(x)x$pc.nt),80,80)))
sh = sh[!is.na(sh)]
sh2=as.numeric(matrix.builder(pc.chart(lapply(MF.x.list$TCRd,function(x)x$pc.nt),80,80)))
sh2 = sh2[!is.na(sh2)]
#E(g)$width = sh/6.25 # proportional to the number of clones sampled in relationship to alpha/beta
E(g)$width = sh # proportional to the number of clones sampled in relationship to alpha/beta

E(g)$color =gg_color_hue(4)[2]
g3 = add_edges(g,c(1,2, 1,3, 1,4, 2,3, 2,4, 3,4))
E(g3)$width[7:12] = sh2
E(g3)$color[7:12] = gg_color_hue(4)[4]
E(g3)$curved = rep(c(-0.15,0.15),each=6)




plot(g3,vertex.frame.color = "grey", vertex.color  = "grey",vertex.size = 20, vertex.label.color = "black",layout = lkk)
}

pub.mean = data.frame(gene = factor(c("TCRg","TCRd"),levels =c("TCRg","TCRd") ),mean = c(mean(sh,na.rm = TRUE),mean(sh2,na.rm = TRUE)))

sh.names = c("2_3","2_4","2_5","3_4","3_5","4_5")
sh.g.df = data.frame(sh,sh.names)
sh.d.df = data.frame(sh2,sh.names)

sh.table.1= cbind(sh.g.df, x = 1)
sh.table.2= cbind(sh.d.df, x = 2)
colnames(sh.table.2)[1] = "sh"
sh.table = rbind(sh.table.1,sh.table.2)
```

```{r Publicity 9, echo = FALSE,fig.width= 10,fig.height=12}

ggplot(data = pub.mean) + geom_bar(mapping = aes (x = gene , y=mean, fill = gene) , width = 0.3,stat = "identity") + 
  ggtitle ("Mean pairwise overlap top 80 clones") + 
  geom_point(data = sh.g.df,mapping = aes (x = 1, y = sh)) +
  geom_line (data = sh.table,mapping = aes(x = x, y = sh,group = sh.names)) +
  geom_text(data = sh.g.df,mapping = aes (x = c(0.8,0.8,0.8,rep(0.8,3)), y = sh, label = sh.names,fontface = "italic"),size = 8) +
  theme(text = element_text(size=20) ) +
  geom_point(data = sh.d.df,mapping = aes (x = 2, y = sh2)) +
  geom_text(data = sh.d.df,mapping = aes (x = c(2.2,2.2,2.4,2.4,rep(2.2,2)), y = sh2+(c(0,0,0,0,0,0)), label = sh.names,fontface = "italic"),size = 8)  +
  scale_fill_manual(values=c( gg_color_hue(4)[2],gg_color_hue(4)[4]))+
  theme(text = element_text(size=20),
        axis.text=element_text(size=20)) 


```

# Length plots


```{r Length plots, echo=FALSE}


ggplot(data = MF.x.df$TCRa[MF.x.df$TCRa$L <60,]) + geom_bar( mapping = aes (x = L),position = "dodge") + labs(title="TCR alpha")
ggplot(data = MF.x.df$TCRa[MF.x.df$TCRa$L <60,]) + geom_bar( mapping = aes (x = L,weight = Umi.count),position = "dodge") + labs(title="TCR alpha weighted")
ggplot(data = MF.x.df$TCRb[MF.x.df$TCRb$L <60,]) + geom_bar( mapping = aes (x = L),position = "dodge") + labs(title="TCR beta")
ggplot(data = MF.x.df$TCRb[MF.x.df$TCRb$L <60,]) + geom_bar( mapping = aes (x = L,weight = Umi.count),position = "dodge") + labs(title="TCR beta weighted")

ggplot(data = MF.x.df$TCRg[MF.x.df$TCRg$L <80,]) + geom_bar( mapping = aes (x = L),position = "dodge") + labs(title="TCR gamma")
ggplot(data = MF.x.df$TCRg[MF.x.df$TCRg$L <80,]) + geom_bar( mapping = aes (x = L,weight = Umi.count),position = "dodge") + labs(title="TCR gamma weighted")
ggplot(data = MF.x.df$TCRd[MF.x.df$TCRd$L <80,]) + geom_bar( mapping = aes (x = L),position = "dodge") + labs(title="TCR delta")
ggplot(data = MF.x.df$TCRd[MF.x.df$TCRd$L <80,]) + geom_bar( mapping = aes (x = L,weight = Umi.count),position = "dodge") + labs(title="TCR delta weighted")
ggplot(data = MF.x.df$IgH[MF.x.df$IgH$L <55,]) + geom_bar( mapping = aes (x = L),position = "dodge")+ labs(title="IgH")
ggplot(data = MF.x.df$IgH[MF.x.df$IgH$L <55,]) + geom_bar( mapping = aes (x = L,weight = Umi.count),position = "dodge")+ labs(title="IgH weighted")

```

# Alpha Delta usage

predominant delta are at least 10 times more frequent molecules as delta than alpha


```{r AD usage plots, echo=FALSE}
load("MF_Vas.5")
MF.ad.usage = cbind(data.frame(as.numeric(table(factor(MF.x.df$TCRa$V.gene,levels =names(MF_Vas.5))))),data.frame(as.numeric(table(factor(MF.x.df$TCRd$V.gene,levels = names(MF_Vas.5))))))
colnames(MF.ad.usage) = c("alphas","deltas")
rownames(MF.ad.usage) = names(MF_Vas.5)
MF.ad.usage = mutate(MF.ad.usage,exclusive = alphas ==0 | deltas ==0 )
MF.ad.usage$type = "predominant alpha"
MF.ad.usage$type[MF.ad.usage$delta ==0 & MF.ad.usage$alpha ==0] = "not used"
MF.ad.usage$type[MF.ad.usage$exclusive == TRUE & MF.ad.usage$delta >0] = "exclusive delta"
MF.ad.usage$type[MF.ad.usage$exclusive == TRUE & MF.ad.usage$alpha >0] = "exclusive alpha"

MF.ad.usage$type[MF.ad.usage$deltas >10 & MF.ad.usage$alpha >0& MF.ad.usage$alpha < 50] = "predominant delta"

MF.ad.usage %>%
  ggplot(mapping = aes(x = alphas, y = deltas, color = type)) +
  geom_point() +
  ggtitle ("TCR alpha vs delta usage for V elements") +
  xlab ("usage as TCR alpha (clones)") +
  ylab ("usage as TCR delta (clones)") 
#kable(MF.ad.usage)
ggplot(MF.ad.usage[MF.ad.usage$alphas>0 & MF.ad.usage$deltas >0,]) + 
  geom_point( mapping = aes(x = 1:length(alphas), y = log10 (alphas/deltas))) + 
  ggtitle ("log ratio of recombination with alpha and delta") +
  xlab ("V element") +
  ylab ("log(ratio a/d)") 

#plot(apply(MF.ad.usage[MF.ad.usage$alphas>0 & MF.ad.usage$deltas >0,],1,function(x)log2(as.numeric(x[1])/as.numeric(x[2]))),ylab = "log2 alpha/delta usage",xlab = "shared v elements")
kable(MF.ad.usage[MF.ad.usage$alphas>0 & MF.ad.usage$deltas >0,])
```

```{r jitter plots, echo=FALSE}
 ggplot(data = MF.x.df$TCRa) + geom_jitter( mapping = aes (x = pub, y = Umi.proportion,color = fish))  +
  ggtitle ("TCR alpha")
 ggplot(data = MF.x.df$TCRb) + geom_jitter( mapping = aes (x = pub, y = Umi.proportion,color = fish)) +
  ggtitle ("TCR beta")
 ggplot(data = MF.x.df$TCRg) + geom_jitter( mapping = aes (x = pub, y = Umi.proportion,color = fish)) +
     ggtitle ("TCR gamma")

 ggplot(data = MF.x.df$TCRd) + geom_jitter( mapping = aes (x = pub, y = Umi.proportion,color = fish)) +
   ggtitle ("TCR delta")
 ggplot(data = MF.x.df$IgH) + geom_jitter( mapping = aes (x = pub, y = Umi.proportion,color = fish)) +
      ggtitle ("Ig H")


```

```{r heatmaps, echo=FALSE}
hm.table = table(MF.x.df$TCRa$V.gene,MF.x.df$TCRa$fish)
hm.table = apply(hm.table,2,function(x)x/sum(x))
hm.counts = as.numeric(hm.table)
hm.df = data.frame(counts = hm.counts,v=rownames(hm.table),fish=rep(MF.fish.names,each=nrow(hm.table)))
ggplot(data = hm.df, mapping = aes(y = v , x = fish, fill = counts )) + geom_tile() + labs(fill = "fraction of UMI counts")


hm.table = table(c(1:61,MF.Ja.alleles.names)[as.numeric(MF.x.df$TCRa$J.gene)],MF.x.df$TCRa$fish)
hm.table = apply(hm.table,2,function(x)x/sum(x))
hm.counts = as.numeric(hm.table)
hm.df = data.frame(counts = hm.counts,j=rownames(hm.table),fish=rep(MF.fish.names,each=nrow(hm.table)))
ggplot(data = hm.df, mapping = aes(y = j , x = fish, fill = counts )) + geom_tile() + labs(title = "tcra assemblies", fill = "fraction of UMI counts")

hm.table = table(MF.x.df$TCRb$V.gene,MF.x.df$TCRb$fish)
hm.table = apply(hm.table,2,function(x)x/sum(x))
hm.counts = as.numeric(hm.table)
hm.df = data.frame(counts = hm.counts,v=rownames(hm.table),fish=rep(MF.fish.names,each=nrow(hm.table)))
ggplot(data = hm.df, mapping = aes(y = v , x = fish, fill = counts )) + geom_tile() + labs(title = "tcrb assemblies", fill = "fraction of UMI counts")

hm.table = table(MF.x.df$TCRb$J.gene,MF.x.df$TCRb$fish)
hm.table = apply(hm.table,2,function(x)x/sum(x))
hm.counts = as.numeric(hm.table)
hm.df = data.frame(counts = hm.counts,j=rownames(hm.table),fish=rep(MF.fish.names,each=nrow(hm.table)))
ggplot(data = hm.df, mapping = aes(y = j , x = fish, fill = counts )) + geom_tile() + labs(title = "tcrb assemblies", fill = "fraction of UMI counts")
```
```